Pool Model

### 2025-08-28 pool_model.tex
adapt to twocolumn layout

@@

    \includegraphics[width=0.6\textwidth]{Figures/TPM_fig.pdf}

@@

\label{eq:reactions_3pools}

@@

    \dot{L} = -(\mu + \lambda) L\\
    \dot{G} = \lambda L + \alpha G - \mu' G\\
    \dot{D} = \mu  L + \mu' G- \beta D  
\label{eq:ode_3pools}

@@

        \dot{L} = -(\mu + \lambda) L\\
        \dot{G} = \lambda L + \alpha R G-\mu' G\\
        \dot{D} = \mu  L + \mu' G- \beta D\\
        \dot{R} = - \frac{\alpha}{N_t} R G

@@

\label{eq:resource}

@@

        \dot{L} = -(\mu + \lambda) L\\
        \dot{G} = \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\mu' G\\
        \dot{D} = \mu  L + \mu' G- \beta D 

@@

    \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource.pdf}

@@

        L(t) = N_0 e^{-(\mu+\lambda)t}\\
        D(t) = N_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}
\label{eq:lagtime_deriv1}

@@

        \dot{G} = \lambda L + \alpha G\left(1-\frac{N}{N_t}\right)\\
        \dot{N} = - \beta D + \alpha G\left(1-\frac{N}{N_t}\right)
    \end{cases}
\label{eq:lagtime_deriv2}

@@

\label{eq:lagtime_deriv3}

@@

\label{eq:lagtime}

@@

@@

    \end{cases}

@@

    \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}

@@

    \dot{L}_A &=& - \lambda_A  L_A + \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right)L_A\\
    \dot{L}_B &=& - \lambda_B L_B + \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right)L_B \\
    \dot{G}_A &=&  \lambda_A  L_A - \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right)L_A +\alpha_A G_A - \frac{\alpha_A}{N_t}\left(L_A+G_A+L_B+G_B\right)G_A\\
    \dot{G}_B &=&  \lambda_B L_B - \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right)L_B  +\alpha_B G_B -\frac{\alpha_B}{N_t}\left(L_A+G_A+L_B+G_B\right)G_B

@@

which has the structure of a \ac{glv} expanded with different internal states for one species.

@@

    \includegraphics[width=0.6\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}

@@

    \dot{G}_A = \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\alpha_A}{N_t}G_AG_B\\
    \dot{G}_B = \alpha_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\alpha_B}{N_t}G_AG_B. 

@@

    \dot{T} &=& \mathcal{F}(R,G).

@@

        \dot{G}_A = \lambda_A R L_A + \alpha_A R G_A - \frac{\mu}{\kappa N_t} T G_A\\
        \dot{G}_B = \lambda_B R L_B + \alpha_B R G_B\\ 
        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
        \dot{T}_{B}^{1,2} = \mathcal{F}^{1,2} (R, G_B) \\

@@

    \mathcal{F}^1(R,G)&=&\kappa G_B\\
    \mathcal{F}^2(R,G)&=&\kappa\alpha_B R G_B.

@@

    \includegraphics[width=1.\textwidth]{Figures/pool_model_2pools_toxin.pdf}

@@

    \dot{G}_A =\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
    \dot{G}_B =\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right). 
    \end{cases}

@@

We can separate three cases (see Figure~\ref{fig:1pool_2sp_inhibit}):

@@

    \psi&>&1:\quad\text{Species B growth faster and/or inhibits A stronger}\\
    \psi&<&1:\quad\text{Species A growth faster and/or inhibits B stronger}

@@

    \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}

@@

    \includegraphics[width=0.5\textwidth]{Figures/pool_model_2sp_1pool_coop.pdf}

@@

    \includegraphics[width=0.4\textwidth]{Figures/snapshot-2-cells-before-division.png}
    \includegraphics[width=0.4\textwidth]{Figures/snapshot-2-cells-after-division.png}

@@

\begin{table}[h]

@@

    \begin{tabularx}{\textwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}

@@

\end{table}

@@

    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00006000.png}\hspace{0.02\textwidth}%
    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00012000.png}\vspace{0.02\textwidth}\\
    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00018000.png}\hspace{0.02\textwidth}%
    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00024000.png}%

@@

    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
        \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}

@@

    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
        \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}

@@

    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
        \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}

@@

    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-random/abm_ode_comparison.png}%
        \subcaption{Randomized initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}

@@

@@

>>>>

\twocolumn

@@

    \includegraphics[width=\columnwidth]{Figures/TPM_fig.pdf}

@@

@@

\begin{align}\begin{split}
    \dot{L} &= -(\mu + \lambda) L\\
    \dot{G} &= \lambda L + \alpha G - \mu' G\\
    \dot{D} &= \mu  L + \mu' G- \beta D  
\end{split}\end{align}.

@@

    \begin{split}
        \dot{L} &= -(\mu + \lambda) L\\
        \dot{G} &= \lambda L + \alpha R G-\mu' G\\
        \dot{D} &= \mu  L + \mu' G- \beta D\\
        \dot{R} &= - \frac{\alpha}{N_t} R G
    \end{split}.

@@

@@

    \begin{split}
        \dot{L} &= -(\mu + \lambda) L\\
        \dot{G} &= \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\mu' G\\
        \dot{D} &= \mu  L + \mu' G- \beta D 
    \end{split}.

@@

    \includegraphics[width=\columnwidth]{Figures/pool_model_3pools_resource.pdf}

@@

    \begin{split}
        L(t) &= N_0 e^{-(\mu+\lambda)t}\\
        D(t) &= N_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}
    \end{split}.

@@

    \begin{split}
        \dot{G} &= \lambda L + \alpha G\left(1-\frac{N}{N_t}\right)\\
        \dot{N} &= - \beta D + \alpha G\left(1-\frac{N}{N_t}\right)
    \end{split}

@@

@@

@@

    \begin{split}

@@

    \end{split}

@@

    \includegraphics[width=\columnwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}

@@

    \dot{L}_A &=& - \lambda_A  L_A + \frac{\lambda_A}{N_t}L_A\left<LG\right>\\
    \dot{L}_B &=& - \lambda_B L_B + \frac{\lambda_B}{N_t}L_B \left<LG\right>\\
    \dot{G}_A &=&  \lambda_A  L_A + \alpha_A G_A - \frac{\lambda_A L_A + \alpha_A G_A}{N_t} \left<LG\right>\\
    \dot{G}_B &=&  \lambda_B L_B - \alpha_B G_B \frac{\lambda_B L_B + \alpha_B G_B}{N_t}\left<LG\right>

@@

where we used the short notation $\left<LG\right>=(L_A+G_A+L_B+G_B)$.
These equations have the structure of a \ac{glv} expanded with different internal states for one species.

@@

    \includegraphics[width=\columnwidth]{Figures/pool_model_2pools_resource_competition.pdf}

@@

    \begin{split}
        \dot{G}_A &= \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\alpha_A}{N_t}G_AG_B\\
        \dot{G}_B &= \alpha_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\alpha_B}{N_t}G_AG_B. 

@@

    \dot{T} = \mathcal{F}(R,G).

@@

    \begin{split}
        \dot{L}_A &= - \lambda_A R L_A\\
        \dot{L}_B &= - \lambda_B R L_B \\
        \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A - \frac{\mu}{\kappa N_t} T G_A\\
        \dot{G}_B &= \lambda_B R L_B + \alpha_B R G_B\\ 
        \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
        \dot{T}_{B}^{1,2} &= \mathcal{F}^{1,2} (R, G_B) \\
    \end{split}.

@@

    \mathcal{F}^1(R,G)&=\kappa G_B\\
    \mathcal{F}^2(R,G)&=\kappa\alpha_B R G_B.

@@

    \includegraphics[width=\columnwidth]{Figures/pool_model_2pools_toxin.pdf}

@@

    \begin{split}
        \dot{G}_A &=\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
        \dot{G}_B &=\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right). 
    \end{split}

@@

We can separate three cases (which are displayed in Figure~\ref{fig:1pool_2sp_inhibit}):

@@

    \psi&>&1:\quad\text{B grows faster/inhibits A}\\
    \psi&<&1:\quad\text{A grows faster/inhibits B}

@@

    % TODO make this into 2x2 grid for plots and insert eqnarray from before as well
    \includegraphics[width=\columnwidth]{Figures/pool_model_1pool_inhib.pdf}

@@

    \includegraphics[width=\columnwidth]{Figures/pool_model_2sp_1pool_coop.pdf}

@@

    \includegraphics[width=0.49\columnwidth]{Figures/snapshot-2-cells-before-division.png}%
    \hspace{0.01\columnwidth}%
    \includegraphics[width=0.49\columnwidth]{Figures/snapshot-2-cells-after-division.png}

@@

\begin{table*}[h]

@@

    \begin{tabularx}{\columnwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}

@@

\end{table*}

@@

    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00006000.png}\hspace{0.01\columnwidth}%
    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00012000.png}\vspace{0.01\columnwidth}\\
    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00018000.png}\hspace{0.01\columnwidth}%
    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00024000.png}%

@@

    \begin{subfigure}[c]{0.5\columnwidth}
        \includegraphics[width=\columnwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
        \subcaption{Homogenous initial cell distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}

@@

    \begin{subfigure}[c]{0.5\columnwidth}
        \includegraphics[width=\columnwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
        \subcaption{Homogenous initial cell distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}

@@

    \begin{subfigure}[c]{0.5\columnwidth}
        \includegraphics[width=\columnwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
        \subcaption{Heterogeneous initial cell distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}

@@

    \begin{subfigure}[c]{0.5\columnwidth}
        \includegraphics[width=\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
        \subcaption{Randomized initial cell distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}

@@

        The diffusion constants are chosen $D_R=D_I=D$.

@@

\onecolumn
